---
title: "quaketools-examples"
author: "danielfs88"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{quaketools-examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

The goal of quakeTools is to clean and plot earthquake data from **The 
Significant Database Earthquake** of [NOAA](https://www.ngdc.noaa.gov/nndc/struts/form?t=101650&s=1&d=1).

## Examples

### First step: extract and clean data

```{r}
library(quakeTools)
library(ggplot2)
cleanquakes <- eq_clean_data(earthquakes)
plotquakes <- dplyr::filter(cleanquakes, COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000)
```

The function `eq_clean_data` receives NOAA dataframe (embedded in this package as `earthquakes`), creating and treating some columns. This function also calls `eq_location_clean` function to treat the `LOCATION_NAME` column specifically, as you could see below.

```{r}
head(plotquakes[,c("DATE", "LATITUDE", "LONGITUDE", "LOCATION")])
```
### First plot example: earthquake timeline

```{r plot_1, , fig.width = 5, fig.height = 3, fig.show='hold'}
ggplot(plotquakes, aes(x = DATE, y = COUNTRY, size = EQ_PRIMARY, colour = DEATHS)) +
geom_timeline() +
xlab("Date") + ggplot2::ylab("") +
labs(size = "Richter scale", color = "# deaths" ) +
theme(axis.line.y=element_blank(),panel.background=element_blank())
```

### Second plot example: earthquake timeline with annotation

```{r plot_2, fig.width = 8, fig.height = 5, fig.show='hold'}
ggplot(plotquakes, aes(x = DATE, y = COUNTRY, size = EQ_PRIMARY, colour = DEATHS)) +
geom_timeline() +
geom_timeline_label(aes(label = LOCATION, n_max = 6)) +
xlab("Date") + ggplot2::ylab("") +
labs(size = "Richter scale", color = "# deaths" ) +
theme(axis.line.y=element_blank(),panel.background=element_blank())
```

### Third plot example: earthquake spots in map with date popup

```{r plot_3, fig.show='hold'}
eq_map(plotquakes, annot_col = "DATE")
```

### Fourth plot example: earthquake spots in map with location, intensity and deaths popup

```{r plot_4, fig.show='hold'}
plotquakes$popup_text = eq_create_label(plotquakes)
eq_map(plotquakes, annot_col = "popup_text")
```
